---
title: 'Statistical Inference course project: Tooth Growth data set exploratory analysis'
author: "Enrico Gaffo"
date: "25 ottobre 2015"
output: pdf_document
---

## Synopsis
We're going to analyze the ToothGrowth data in the R datasets package.

## Exploratory Analysis
Load the ToothGrowth data and perform some basic exploratory data analyses 
Provide a basic summary of the data.

```{r, echo=TRUE}
library(datasets)
data("ToothGrowth")
summary(ToothGrowth)
library(GGally)
ggpairs(ToothGrowth)
library(ggplot2)
qplot(factor(dose), len, fill = supp, data = ToothGrowth, 
      geom = "boxplot", position = "dodge") + geom_jitter() +
  xlab("Vitamin C dose") + 
  ylab("length of odontoblasts (teeth)")
```

Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

```{r, echo=TRUE}
table(ToothGrowth$supp, ToothGrowth$dose)
library(reshape2)
wtgs <- split(ToothGrowth, ToothGrowth$supp)
t.test(wtg$OJ$len, wtg$VC$len, paired = F)
# wtgd <- split(ToothGrowth, ToothGrowth$dose)
# t.test(len~dose, data = ToothGrowth, paired = F)
```


